Optical characterization of gaps in directly bonded Si compound optics using infrared spectroscopy

### 2014-11-26 AO_bondedSi.tex
fix tons of broken references, what a waste of time

\subsection{Why Silicon is good}
Crystalline silicon is an excellent optical material for the infrared. Si has a high refractive index ($\sim$3.4, Frey et al. 2006) and superb transmission from slightly longer than the equivalent wavelength of its band gap ($\sim$1.15 $\mu$m) to about 6.5 $\mu$m \cite{PhysRev.108.268, PhysRev.78.178} and in the far-IR \cite{doi:10.1117/12.323764}.  It has adequate transmission  ($\alpha \sim 1$ cm$^{-1}$) for many purposes in the 18-30 $\mu$m region as well \cite{doi:10.1117/12.323764}.  There is an extensive industrial infrastructure for the production of ultrapure Si and for polishing and lithographic patterning and etching of optical parts.

\subsection{Why bonding is good and AR Coating is bad}
As with a number of glasses and crystalline optical materials, it is possible to form strong physical bonds between two planar or conformal silicon surfaces.  One advantage of this direct Si-Si bonding of macroscopic optical parts is that it facilitates the manufacture of complex optical subsystems, for example double-convex aspheres, combinations of lenses, prisms, and transmission gratings, and pairs of gratings oriented at right angles \cite{2012SPIE.8450E..2TV, 2010SPIE.7739E.123G}.  A second important advantage is that a sufficiently close bond is optically lossless, an important consideration not only for its effect on throughput but also because it can eliminate concerns about optical ghosts.  The remaining advantages of bonding stem from various difficulties encountered in antireflection coating a high index material in the near and mid-infrared:  There is a limited choice of index-matching materials, in particular at longer wavelengths.  Systems often are used at cryogenic temperatures and considerations of differential thermal expansion, brittleness and hygroscopic properties can restrict the choices even further. It is hard to get good antireflection performance across broad spectral ranges and at a range of incidence angles.

@@

There are several good reviews of Si bonding that include accounts of the history of the field and a description of process alternatives, metrology, and problems (Goesele and Tong, Mastreika et al.  2014).  The foundation paper on Si bonding came in 1986[14]. After that, there was a sequence of papers that used IR imaging to look for evidence of defects in the bond. Stengl et al. (1988) and Goesele et al. (1995) [23, 24] directly monitored the bond propagation with IR imaging. Lehman et al. 1989 and Goesele et al. [19, 23] demonstrated gap-free bonding in a micro- cleanroom, in which the bonding process occurs immediately after a rinse-spin-vacuum cycle inside of a vacuum enclosure.  Other authors have investigated the growth of bubbles at bond with time and the elimination of some or all of these bubbles after annealing bonded parts at temperatures of 800-1100 C (REFS including looking at some from the Mastreika paper).

The chemistry of the bonding interface began to be understood with infrared absorption spectroscopy and multiple internal reflection spectroscopy[25], which lent credibility to the idea of distinct temperature phases and evidence for Si?H bonds[26]. A robust picture of the chemistry as a function of anneal temperature was ultimately fleshed out[27, 28]. Other variables potentially affecting bond strength were studied, for example surface roughness[29], surface topography [30], surface preparation[31], annealing time [32], ambient pressure and substrate thickness [33, 34], and moisture and defects [35]. Notably, [29] found that surface roughness above 1.3 nm RMS results in poor bonding quality in their sample of wafers etched with Argon. Results from studies of surface topography are inconclusive[30].

@@

If Si-Si bonds are to be a useful part of the optical designer's toolbox, the bonds must have transmission losses similar to or lower than those of two antireflection coated surfaces over the operating wavelength (typically $< 2$\%).  If the optical layout does not eliminate pupil ghosts from the flat Si surfaces, there are additional requirements on the reflectivity of individual bubbles in the bond. While completely bonded Si parts will be completely lossless at the interface, problems on different scales with different physical origins can lead to less-than-perfect bonds.  On the smallest scales, imperfect cleaning can leave small voids where particulates hold the bonding planes apart or where hydrocarbons can serve as catalysts for gap formation[17].  Even in the best of cleanrooms, particles at the optically significant 10-20 nm scale can be present in relatively high abundance. Figure 5 of Cooper 1986[16] shows the surface flux (particles cm$^2$ s$^{-1}$) vs. particle size ($\mu$m). There will be roughly 104 times more 10 nm sized particles than 1 ?m sized particles.  On larger scale, gaseous hydrogen, which is a byproduct of the  hydrophilic bonding process, can migrate into the bond and form local bubbles (REFS).  Most of the bonding literature concerns itself with wafer bonding.  In bonding wafers to each other or to thicker optically polished Si disks, one or both of the partners can conform as the bonding front propagates.  When bonding two thicker and more rigid substrates, differences in the shapes of the polished surfaces can potentially lead to topological inconsistencies that prohibit some areas from bonding.

\subsection{Existing detection schemes}
Verification that an optical bond meets its requirements and possible diagnosis of problems in the bonding technique require an accurate assessment of the geometry and lateral and axial dimensions of any defects.  Since silicon is opaque in the optical, the most straightforward technique for detecting defects in the bond is to image it in the infrared.  Fresnel losses will cause unbonded regions to appear darker than regions where the bond is perfect and the transmission is 100\%.  In cases where the axial extent of the gap is a wavelength or more, Newton rings appear in the image of the gap [20].  X-ray topography [18, 22] exploits the change in optical density along lines of sight to allow precise high spatial resolution measurements of the spatial extent of gaps with axial dimensions as small as a few nanometers but not the axial extent itself. In ultrasound microscopy, the density change at the gap interface produces reflections of sound waves and enables mapping the lateral gap dimensions but produces limited information about the axial dimension [21].

@@

\section{History of Si direct bonding}
\label{secHistory}

The foundation paper on Si bonding came in 1986\cite{1986JAP....60.2987S}.  After that, there was a sequence of papers using IR imaging to look for evidence of defects in the bond.  \cite{1988JaJAP..27L2364S, 1995ApPhL..67.3614G} directly monitored the bond propagation with IR imaging.  \cite{1988JaJAP..27L2364S, 1989JaJAP..28L2141L} demonstrated gap-free bonding in a micro-cleanroom, in which the bonding process occurs immediately after a rinse-spin-vacuum cycle inside of a vacuum enclosure.  Our group has not attempted the micro-cleanroom technique owing to the challenge of spin-cleaning large (30 mm thick) Si substrates.  It is not clear whether the custom micro-cleanroom technique has gained wide usage.  \cite{1989JaJAP..28L2141L} directly measured the bond surface energy.  \cite{Mitani1990} discussed the causes and remedies for bubble formation.  Specifically, the IR bubble density changes as a function of anneal temperature\cite{1992JEMat..21..669M}, with bubbles disappearing at temperatures greater than 1100$^\circ$ C.  The chemistry of the bonding interface began to be understood with infrared absorption spectroscopy and multiple internal reflection spectroscopy\cite{feijoo1994}, which lent credibility to the idea of distinct temperature phases and evidence for Si$-$H bonds\cite{1995ApPhA..61..101R}.  A robust picture of the chemistry as a function of anneal temperature was ultimately fleshed out\cite{1996JaJAP..35.2102R, 1998AnRMS..28..215G}.  Other variables potentially affecting bond strength were studied, for example surface roughness\cite{JJAP.37.4197}, surface topography \cite{2001JOptA...3...85G}, surface preparation\cite{1996ApPhL..68.2222T}, annealing time \cite{2000JAP....88.4404H}, ambient pressure and substrate thickness \cite{1995ApPhL..67..863G, 2007ApOpt..46.6793H}, and moisture and defects \cite{2001JAP....89.6013L}.  Notably, \cite{JJAP.37.4197} found that surface roughness above 1.3 nm RMS results in poor bonding quality in their sample of wafers etched with Argon.  Results from studies of surface topography are inconclusive\cite{2001JOptA...3...85G}.  Nowhere in the literature did we find a technique similar to the one presented here, probably because the technique relies on the high precision of relatively new IR capable spectroscopy that was not widely available until this decade.

@@

It is conceivable that there are particles in between the Si interfaces.  Particle contaminants are discussed in \cite{Mitani1990}.

@@

Our measurement technique exploits the large Fresnel reflection \cite{2001opt4.book.....H} at the Si$-$vacuum interface.  The refractive index of silicon at room temperature ranges from 3.55$-$3.45 from $\lambda = $ 1150$-$2500 nm \cite{2006SPIE.6273E..77F}, one of the largest refractive indices of any conventional material.  The Fresnel reflection is about 30\% per surface.  For comparison, conventional optical materials with refractive indices in the range 1.5$-$1.7 have Fresnel reflection of merely 4$-$7\% per surface.  Figure \ref{figSiIndexFinesse} shows the refractive index as a function of wavelength for room temperature, which we computed using the coefficients from \cite{2006SPIE.6273E..77F}.

\begin{figure}[htbp]
\centerline{\includegraphics[width=0.95\columnwidth]{figs/SiIndexAOmgsFinesseFig.pdf}}
\caption{Plot of refractive index, $n$, and coefficient of finesse, $F$, as a function of wavelength, $\lambda$.\label{figSiIndexFinesse} We computed the temperature dependent refractive index from the tabulations of \cite{2006SPIE.6273E..77F}.  The coefficient of finesse is defined in Equation \ref{eq:FabPerot}.  The minuscule temperature dependence is largely negligible, and for the rest of the paper we assume room temperature refractive index.}
\end{figure}

The Si$-$Si interface gap is a low-Finesse Fabry-P\`{e}rot etalon\cite{2007fuph.book.....S}.  The transmission through a Fabry-P\`{e}rot depends on the wavelength of light, the reflectivity of the etalon sidewalls, and the size of the gap.  We assume transmission is at normal incidence, the refractive index of the gap is 1.0, and the material is at room temperature.  We compute the reflectivity for an Si$-$vacuum interface, which is a modest function of wavelength through the refractive index of Silicon, $n(\lambda, T)$ \cite{2006SPIE.6273E..77F}, as shown in Figure \ref{figSiIndexFinesse}.  In the equations ahead we drop the subscripts on the refractive index $n(\lambda, T) \rightarrow n$ and the coefficient of finesse $F(\lambda, T) \rightarrow F$ for clarity.  The computation is simply Fresnel's law at normal incidence:

@@

The reflectivity is about 30\% in the wavelength range 1100 to 5000 nm.  We've defined the coefficient of finesse\cite{2007fuph.book.....S}, $F$, a customary way to encapsulate the Fabry-P\`{e}rot etalon's dependence on the reflectivity.  The coefficient of finesse is shown in the right panel of Figure \ref{figSiIndexFinesse}.  Si is negligibly absorbing longward of about $\lambda$ = 1200 nm, which we verified by comparing the transmission of Si reference samples of different thicknesses, as shown in Figure \ref{figSiAbsorbfig}.  Figure \ref{figSiAbsorbfig} also verifies a key assumption that we will make when comparing the transmission of Si substrates of different heritage and thickness.  Specifically DSP Si pucks and wafers have the indistinguishable transmission longward of 1250 nm.  The Si refractive index is indistinguishable from sample to sample for the wavelength ranges we care about- the non-absorbing wavelengths greater than about 1200 nm.  \cite{2006SPIE.6273E..77F} cites the wide variety of values for refractive index from the literature as evidence that batch-specific Si should be used as reference for cases in which absolute accuracy greater than $\pm5\times10^{-3}$ is desired.  An absolute deviation of $\pm5\times10^{-3}$ in refractive index corresponds to a Fresnel transmission difference of about 0.2\%, which is comparable to our measurement uncertainty.  In the next subsection we work out the wavelength dependent transmission for the air etalon.

@@

\subsection{Predicted transmission spectrum for bonded Si with finite interface gap}
\label{secTheory}
There is a key hurdle in computing the transmission spectrum for bonded Si with a finite interface gap- multiple incoherent reflections occur within the Si substrate at normal incidence.  Treatments for multiple coherent reflections have been described in detail in the optics literature \cite{2007fuph.book.....S}.  The wave transfer matrix technique treats each dielectric interface as a matrix with elements relating the pre- and post- interface complex amplitudes in the left and right directions.  We adapted the wave transfer technique for incoherent interactions \cite{2002ApOpt..41.3978K}.  Specifically I constructed an incoherent wave transfer matrix whose elements relate the intensities (and not complex amplitudes) before and after an interface.  The details are worked out in Appendix \ref{sec:Append-IMRTMM}.  We computed the transmission for two scenarios- first a double side polished (DSP) Si sample with no gap, and second a pair of bonded Si samples with a gap thickness $d$.  The DSP Si sample with no air gap has a transmission equal to:

@@

The substrate thicknesses ranged from 0.5 to 3.3 mm per substrate.  We classify substrates with thicknesses larger than 1 mm as ``pucks''.  Substrates with thickness less than 1.0 mm thick are called ``wafers''.  The distinction between the two classes of substrate thickness is based on the anticipation that thin substrates will conform more easily to their bonding partner substrates.  There is also evidence that bond front propagation is slower in thick pucks \cite{2007ApOpt..46.6793H} than in wafers.  Our ultimate goal is to bond pucks with thicknesses as large as 30 mm thick.  These pucks truly resemble over-sized hockey pucks.

@@

We looked for the indication of large gaps detectable as ``IR bubbles'' \cite{1992JEMat..21..669M} with IR imaging.  The detector was an IR Vista $\alpha$NIR infrared focal plane array, with 312 $\times$ 252 pixels, with 30 $\mu$m square pixels.  We used a 50 mm F/2 lens, with a fiber optic white light illumination source.  We estimate that the lamp-detector combination was sensitive to wavelengths in the range 1.1-1.6 $\mu$m.  The field of view sub-sampled the 4 inch diameter wafers, so we dithered the sample.  No effort was made to flat field the detector, so copious detector non-uniformities and vignetting are present in our images.  It was easy to detect IR bubbles despite the detector artifacts since the bubbles translate with the wafer during dithering.  We detected IR bubbles in all 3 of the bonded Si samples on which we recorded IR images.  The bubble areal density varied from about 4 per 100 mm diameter bonded wafer pair to over 20 per bonded wafer pair.  Naturally, the unbonded DSP wafers show no such IR bubbles, which are entirely attributable to the gaps between bonded substrates as discussed in the literature (see Section \ref{secHistory}).  No attempt at quantitative measurement was made on the IR images due to evidence that the detector is non-linear.  The morphology was clear regardless of the detector response.

@@

\subsection{Mitigation of measurement artifacts}
We mitigated measurement artifacts in the following ways.  For the Sharpie marker behaving as an anti-reflection (AR) coating, we simply removed the sharpie with IPA.  We have mostly been able to rule out wavelength-dependent diffraction from the mesh pattern as a measurement artifact, since the beam size is large and the angular spread of diffraction from the mesh pattern would be much smaller than the spread of the beam.  Still, to remove this concern, we moved to using the coarse mesh rather than the medium or fine meshes, when possible.  We saw no difference in the bond quality surrounding the medium or fine meshes, which is consistent with results from the literature \cite{1992JEMat..21..669M}.  Specifically, the gas pressure of bonded wafers grows as the fill factor of gaps decreased.  In other words, what matters is the fill factor of gaps, not their relative sizes.  Since the fine, medium, and coarse meshes had the same fill factors, we anticipate no difference in the pressure of the gaps, and therefore no difference in the bonding properties.

@@

The wave transfer matrix method is described in detail in chapter 7 by Saleh and Teich in ``Fundamentals of Photonics'' \cite{2007fuph.book.....S}.  Briefly, their technique is to assemble a 2$\times$2 scattering matrix $\boldsymbol{S}$ which has the elements:
\begin{eqnarray}
\boldsymbol{S} = \left(
\begin{array}{cc}
 t_{12} & r_{21} \\
 r_{12} & t_{21} \\
\end{array}
\right)
\end{eqnarray}
Where $t$ and $r$ stand for transmission and reflection respectively.  The order of subscripts is the order of origin and destination of the wave in regards to the interface, so we know something about the direction of the wave and how it got there.  The $\boldsymbol{S}$ matrix encapsulates all the information we need to know about how light waves interact with the interface.  The real power of the matrix technique comes from a cousin of the scattering matrix, the so-called wave transfer matrix, $\boldsymbol{M}$.  The matrix $\boldsymbol{M}$ has the convenient property that its output can be used as the input for another matrix.  In other words, the input vector to $\boldsymbol{M}$ is made of the left and right moving components directly before the interface; the outputs are the left and right moving components directly after the interface:
\begin{eqnarray}
\left(
\begin{array}{c}
 U_2^{(+)} \\
 U_1^{(-)} \\
\end{array}
\right)=\boldsymbol{S} \left(
\begin{array}{c}
 U_1^{(+)} \\
 U_2^{(-)} \\
\end{array}
\right) \\
\left(
\begin{array}{c}
 U_2^{(+)} \\
 U_2^{(-)} \\
\end{array}
\right)=\boldsymbol{M} \left(
\begin{array}{c}
 U_1^{(+)} \\
 U_1^{(-)} \\
\end{array}
\right)
\end{eqnarray}

The elements of $\boldsymbol{S}$ and $\boldsymbol{M}$ are related to each other by geometric transformations\cite{2007fuph.book.....S}.  Much of the literature on the transfer matrix method deals with applications related to thin films, in which the wavelength is comparable to the size of the dielectric layer.  For thin films the vector components $U_{i}$ represent the complex amplitudes of the electromagnetic waves.  The polarization state can be encapsulated in scattering matrix components\cite{2007fuph.book.....S}.  The intensities of the emergent spectrum can be computed from the absolute square of the complex amplitudes.  For thick films, the vector components are the intensities of the emergent spectrum, since waves are incoherent.  \cite{2002ApOpt..41.3978K} work out the general transfer-matrix method for optical multilayer systems with incoherent interference.  The key idea for the incoherent transfer matrix method approach is to populate a scattering matrix with elements equal to the (wavelength dependent) transmitted $T_i$ and reflected $R_i$ intensities of an interface or set of interfaces that act together.  Then, use the geometric transformations to construct the $\boldsymbol{M}$ matrix.

@@

From the matrix transformation equations \cite{2007fuph.book.....S} we know $M_{22}=1/T$.  Taking the inverse of the bottom right element of $\boldsymbol{M_{net}}$, we get the transmission $T_{net}$ through the net optical device:

>>>>

Placeholder citation so the document will compile: see: \cite{2012SPIE.8450E..2TV, 2010SPIE.7739E.123G}.

\subsection{Why Silicon is good}
Crystalline silicon is an excellent optical material for the infrared. Si has a high refractive index ($\sim$3.4, Frey et al. 2006) and superb transmission from slightly longer than the equivalent wavelength of its band gap ($\sim$1.15 $\mu$m) to about 6.5 $\mu$m \verb|\cite{PhysRev.108.268, PhysRev.78.178}| and in the far-IR \verb|\cite{doi:10.1117/12.323764}|.  It has adequate transmission  ($\alpha \sim 1$ cm$^{-1}$) for many purposes in the 18-30 $\mu$m region as well \verb|\cite{doi:10.1117/12.323764}|.  There is an extensive industrial infrastructure for the production of ultrapure Si and for polishing and lithographic patterning and etching of optical parts.

\subsection{Why bonding is good and AR Coating is bad}
As with a number of glasses and crystalline optical materials, it is possible to form strong physical bonds between two planar or conformal silicon surfaces.  One advantage of this direct Si-Si bonding of macroscopic optical parts is that it facilitates the manufacture of complex optical subsystems, for example double-convex aspheres, combinations of lenses, prisms, and transmission gratings, and pairs of gratings oriented at right angles \verb|\cite{2012SPIE.8450E..2TV, 2010SPIE.7739E.123G}|.  A second important advantage is that a sufficiently close bond is optically lossless, an important consideration not only for its effect on throughput but also because it can eliminate concerns about optical ghosts.  The remaining advantages of bonding stem from various difficulties encountered in antireflection coating a high index material in the near and mid-infrared:  There is a limited choice of index-matching materials, in particular at longer wavelengths.  Systems often are used at cryogenic temperatures and considerations of differential thermal expansion, brittleness and hygroscopic properties can restrict the choices even further. It is hard to get good antireflection performance across broad spectral ranges and at a range of incidence angles.

@@

There are several good reviews of Si bonding that include accounts of the history of the field and a description of process alternatives, metrology, and problems \verb|\cite{1998AnRMS..28..215G,Masteika2014}|.  The foundation paper on Si bonding came in 1986 \verb|\cite|{1986JAP....60.2987S}|. After that, there was a sequence of papers that used IR imaging to look for evidence of defects in the bond. Stengl et al. (1988) and Goesele et al. (1995) \verb|\cite{1988JaJAP..27L2364S, 1995ApPhL..67.3614G}| directly monitored the bond propagation with IR imaging. Lehman et al. 1989 and Goesele et al. \verb|\cite{1989JaJAP..28L2141L, 1995ApPhL..67.3614G}| demonstrated gap-free bonding in a micro- cleanroom, in which the bonding process occurs immediately after a rinse-spin-vacuum cycle inside of a vacuum enclosure.  Other authors have investigated the growth of bubbles at bond with time and the elimination of some or all of these bubbles after annealing bonded parts at temperatures of 800-1100 $^\circ$C \verb|\cite{Horn2009, Masteika2014}|.

The chemistry of the bonding interface began to be understood with infrared absorption spectroscopy and multiple internal reflection spectroscopy\cite{feijoo1994}, which lent credibility to the idea of distinct temperature phases and evidence for Si$-$H bonds\cite{1995ApPhA..61..101R}.  A robust picture of the chemistry as a function of anneal temperature was ultimately fleshed out\cite{1996JaJAP..35.2102R, 1998AnRMS..28..215G}.  Other variables potentially affecting bond strength were studied, for example surface roughness\cite{JJAP.37.4197}, surface topography \cite{2001JOptA...3...85G}, surface preparation\cite{1996ApPhL..68.2222T}, annealing time \cite{2000JAP....88.4404H}, ambient pressure and substrate thickness \cite{1995ApPhL..67..863G, 2007ApOpt..46.6793H}, and moisture and defects \cite{2001JAP....89.6013L}.  Notably, \cite{JJAP.37.4197} found that surface roughness above 1.3 nm RMS results in poor bonding quality in their sample of wafers etched with Argon.  Results from studies of surface topography are inconclusive\cite{2001JOptA...3...85G}.

@@

If Si-Si bonds are to be a useful part of the optical designer's toolbox, the bonds must have transmission losses similar to or lower than those of two antireflection coated surfaces over the operating wavelength (typically $< 2$\%).  If the optical layout does not eliminate pupil ghosts from the flat Si surfaces, there are additional requirements on the reflectivity of individual bubbles in the bond. While completely bonded Si parts will be completely lossless at the interface, problems on different scales with different physical origins can lead to less-than-perfect bonds.  On the smallest scales, imperfect cleaning can leave small voids where particulates hold the bonding planes apart or where hydrocarbons can serve as catalysts for gap formation.  Even in the best of cleanrooms, particles at the optically significant 10-20 nm scale can be present in relatively high abundance. Figure 5 of Cooper 1986\verb|\cite{doi:10.1080/02786828608959094}| shows the surface flux (particles cm$^2$ s$^{-1}$) vs. particle size ($\mu$m). There will be roughly 104 times more 10 nm sized particles than 1 $\mu$m sized particles.  On larger scale, gaseous hydrogen, which is a byproduct of the hydrophilic bonding process, can migrate into the bond and form local bubbles \verb|\cite{Masteika2014}|.  Most of the bonding literature concerns itself with wafer bonding.  In bonding wafers to each other or to thicker optically polished Si disks, one or both of the partners can conform as the bonding front propagates.  When bonding two thicker and more rigid substrates, differences in the shapes of the polished surfaces can potentially lead to topological inconsistencies that prohibit some areas from bonding.

\subsection{Existing detection schemes}
Verification that an optical bond meets its requirements and possible diagnosis of problems in the bonding technique require an accurate assessment of the geometry and lateral and axial dimensions of any defects.  Since silicon is opaque in the optical, the most straightforward technique for detecting defects in the bond is to image it in the infrared.  Fresnel losses will cause unbonded regions to appear darker than regions where the bond is perfect and the transmission is 100\%.  In cases where the axial extent of the gap is a wavelength or more, Newton rings appear in the image of the gap.  X-ray topography \verb|\cite{Mitani1990}| exploits the change in optical density along lines of sight to allow precise high spatial resolution measurements of the spatial extent of gaps with axial dimensions as small as a few nanometers but not the axial extent itself. In ultrasound microscopy, the density change at the gap interface produces reflections of sound waves and enables mapping the lateral gap dimensions but produces limited information about the axial dimension \verb|\cite{2000RScI...71.1869G}|.

@@

@@

It is conceivable that there are particles in between the Si interfaces.  Particle contaminants are discussed in \verb|\cite{Mitani1990}|.

@@

Our measurement technique exploits the large Fresnel reflection \verb|\cite{2001opt4.book.....H}| at the Si$-$vacuum interface.  The refractive index of silicon at room temperature ranges from 3.55$-$3.45 from $\lambda = $ 1150$-$2500 nm \verb|\cite{2006SPIE.6273E..77F}|, one of the largest refractive indices of any conventional material.  The Fresnel reflection is about 30\% per surface.  For comparison, conventional optical materials with refractive indices in the range 1.5$-$1.7 have Fresnel reflection of merely 4$-$7\% per surface.  Figure \ref{figSiIndexFinesse} shows the refractive index as a function of wavelength for room temperature, which we computed using the coefficients from \verb|\cite{2006SPIE.6273E..77F}|.

\begin{figure}[htbp]
\centerline{\includegraphics[width=0.95\columnwidth]{figs/SiIndexAOmgsFinesseFig.pdf}}
\caption{Plot of refractive index, $n$, and coefficient of finesse, $F$, as a function of wavelength, $\lambda$.\label{figSiIndexFinesse} We computed the temperature dependent refractive index from the tabulations of CITE\texttt{2006SPIE.6273E..77F}.  The coefficient of finesse is defined in Equation \ref{eq:FabPerot}.  The minuscule temperature dependence is largely negligible, and for the rest of the paper we assume room temperature refractive index.}
\end{figure}

The Si$-$Si interface gap is a low-Finesse Fabry-P\`{e}rot etalon\verb|\cite{2007fuph.book.....S}|.  The transmission through a Fabry-P\`{e}rot depends on the wavelength of light, the reflectivity of the etalon sidewalls, and the size of the gap.  We assume transmission is at normal incidence, the refractive index of the gap is 1.0, and the material is at room temperature.  We compute the reflectivity for an Si$-$vacuum interface, which is a modest function of wavelength through the refractive index of Silicon, $n(\lambda, T)$ \verb|\cite{2006SPIE.6273E..77F}|, as shown in Figure \ref{figSiIndexFinesse}.  In the equations ahead we drop the subscripts on the refractive index $n(\lambda, T) \rightarrow n$ and the coefficient of finesse $F(\lambda, T) \rightarrow F$ for clarity.  The computation is simply Fresnel's law at normal incidence:

@@

The reflectivity is about 30\% in the wavelength range 1100 to 5000 nm.  We've defined the coefficient of finesse\verb|\cite{2007fuph.book.....S}|, $F$, a customary way to encapsulate the Fabry-P\`{e}rot etalon's dependence on the reflectivity.  The coefficient of finesse is shown in the right panel of Figure \ref{figSiIndexFinesse}.  Si is negligibly absorbing longward of about $\lambda$ = 1200 nm, which we verified by comparing the transmission of Si reference samples of different thicknesses, as shown in Figure \ref{figSiAbsorbfig}.  Figure \ref{figSiAbsorbfig} also verifies a key assumption that we will make when comparing the transmission of Si substrates of different heritage and thickness.  Specifically DSP Si pucks and wafers have the indistinguishable transmission longward of 1250 nm.  The Si refractive index is indistinguishable from sample to sample for the wavelength ranges we care about- the non-absorbing wavelengths greater than about 1200 nm.  \verb|\cite{2006SPIE.6273E..77F}| cites the wide variety of values for refractive index from the literature as evidence that batch-specific Si should be used as reference for cases in which absolute accuracy greater than $\pm5\times10^{-3}$ is desired.  An absolute deviation of $\pm5\times10^{-3}$ in refractive index corresponds to a Fresnel transmission difference of about 0.2\%, which is comparable to our measurement uncertainty.  In the next subsection we work out the wavelength dependent transmission for the air etalon.

@@

\subsection{Predicted transmission spectrum for bonded Si with finite interface gap}
\label{secTheory}
There is a key hurdle in computing the transmission spectrum for bonded Si with a finite interface gap- multiple incoherent reflections occur within the Si substrate at normal incidence.  Treatments for multiple coherent reflections have been described in detail in the optics literature \verb|\cite{2007fuph.book.....S}|.  The wave transfer matrix technique treats each dielectric interface as a matrix with elements relating the pre- and post- interface complex amplitudes in the left and right directions.  We adapted the wave transfer technique for incoherent interactions \verb|\cite{2002ApOpt..41.3978K}|.  Specifically I constructed an incoherent wave transfer matrix whose elements relate the intensities (and not complex amplitudes) before and after an interface.  The details are worked out in Appendix \ref{sec:Append-IMRTMM}.  We computed the transmission for two scenarios- first a double side polished (DSP) Si sample with no gap, and second a pair of bonded Si samples with a gap thickness $d$.  The DSP Si sample with no air gap has a transmission equal to:

@@

The substrate thicknesses ranged from 0.5 to 3.3 mm per substrate.  We classify substrates with thicknesses larger than 1 mm as ``pucks''.  Substrates with thickness less than 1.0 mm thick are called ``wafers''.  The distinction between the two classes of substrate thickness is based on the anticipation that thin substrates will conform more easily to their bonding partner substrates.  There is also evidence that bond front propagation is slower in thick pucks \verb|\cite{2007ApOpt..46.6793H}| than in wafers.  Our ultimate goal is to bond pucks with thicknesses as large as 30 mm thick.  These pucks truly resemble over-sized hockey pucks.

@@

We looked for the indication of large gaps detectable as ``IR bubbles'' \verb|\cite{1992JEMat..21..669M}| with IR imaging.  The detector was an IR Vista $\alpha$NIR infrared focal plane array, with 312 $\times$ 252 pixels, with 30 $\mu$m square pixels.  We used a 50 mm F/2 lens, with a fiber optic white light illumination source.  We estimate that the lamp-detector combination was sensitive to wavelengths in the range 1.1-1.6 $\mu$m.  The field of view sub-sampled the 4 inch diameter wafers, so we dithered the sample.  No effort was made to flat field the detector, so copious detector non-uniformities and vignetting are present in our images.  It was easy to detect IR bubbles despite the detector artifacts since the bubbles translate with the wafer during dithering.  We detected IR bubbles in all 3 of the bonded Si samples on which we recorded IR images.  The bubble areal density varied from about 4 per 100 mm diameter bonded wafer pair to over 20 per bonded wafer pair.  Naturally, the unbonded DSP wafers show no such IR bubbles, which are entirely attributable to the gaps between bonded substrates as discussed in the literature (see Section \ref{secHistory}).  No attempt at quantitative measurement was made on the IR images due to evidence that the detector is non-linear.  The morphology was clear regardless of the detector response.

@@

\subsection{Mitigation of measurement artifacts}
We mitigated measurement artifacts in the following ways.  For the Sharpie marker behaving as an anti-reflection (AR) coating, we simply removed the sharpie with IPA.  We have mostly been able to rule out wavelength-dependent diffraction from the mesh pattern as a measurement artifact, since the beam size is large and the angular spread of diffraction from the mesh pattern would be much smaller than the spread of the beam.  Still, to remove this concern, we moved to using the coarse mesh rather than the medium or fine meshes, when possible.  We saw no difference in the bond quality surrounding the medium or fine meshes, which is consistent with results from the literature \verb|\cite{1992JEMat..21..669M}|.  Specifically, the gas pressure of bonded wafers grows as the fill factor of gaps decreased.  In other words, what matters is the fill factor of gaps, not their relative sizes.  Since the fine, medium, and coarse meshes had the same fill factors, we anticipate no difference in the pressure of the gaps, and therefore no difference in the bonding properties.

@@

The wave transfer matrix method is described in detail in chapter 7 by Saleh and Teich in ``Fundamentals of Photonics'' \verb|\cite{2007fuph.book.....S}|.  Briefly, their technique is to assemble a 2$\times$2 scattering matrix $\boldsymbol{S}$ which has the elements:
\begin{eqnarray}
\boldsymbol{S} = \left(
\begin{array}{cc}
 t_{12} & r_{21} \\
 r_{12} & t_{21} \\
\end{array}
\right)
\end{eqnarray}
Where $t$ and $r$ stand for transmission and reflection respectively.  The order of subscripts is the order of origin and destination of the wave in regards to the interface, so we know something about the direction of the wave and how it got there.  The $\boldsymbol{S}$ matrix encapsulates all the information we need to know about how light waves interact with the interface.  The real power of the matrix technique comes from a cousin of the scattering matrix, the so-called wave transfer matrix, $\boldsymbol{M}$.  The matrix $\boldsymbol{M}$ has the convenient property that its output can be used as the input for another matrix.  In other words, the input vector to $\boldsymbol{M}$ is made of the left and right moving components directly before the interface; the outputs are the left and right moving components directly after the interface:
\begin{eqnarray}
\left(
\begin{array}{c}
 U_2^{(+)} \\
 U_1^{(-)} \\
\end{array}
\right)=\boldsymbol{S} \left(
\begin{array}{c}
 U_1^{(+)} \\
 U_2^{(-)} \\
\end{array}
\right) \\
\left(
\begin{array}{c}
 U_2^{(+)} \\
 U_2^{(-)} \\
\end{array}
\right)=\boldsymbol{M} \left(
\begin{array}{c}
 U_1^{(+)} \\
 U_1^{(-)} \\
\end{array}
\right)
\end{eqnarray}

The elements of $\boldsymbol{S}$ and $\boldsymbol{M}$ are related to each other by geometric transformations\verb|\cite{2007fuph.book.....S}|.  Much of the literature on the transfer matrix method deals with applications related to thin films, in which the wavelength is comparable to the size of the dielectric layer.  For thin films the vector components $U_{i}$ represent the complex amplitudes of the electromagnetic waves.  The polarization state can be encapsulated in scattering matrix components\verb|\cite{2007fuph.book.....S}|.  The intensities of the emergent spectrum can be computed from the absolute square of the complex amplitudes.  For thick films, the vector components are the intensities of the emergent spectrum, since waves are incoherent.  \verb|\cite{2002ApOpt..41.3978K}| work out the general transfer-matrix method for optical multilayer systems with incoherent interference.  The key idea for the incoherent transfer matrix method approach is to populate a scattering matrix with elements equal to the (wavelength dependent) transmitted $T_i$ and reflected $R_i$ intensities of an interface or set of interfaces that act together.  Then, use the geometric transformations to construct the $\boldsymbol{M}$ matrix.

@@

From the matrix transformation equations \verb|\cite{2007fuph.book.....S}| we know $M_{22}=1/T$.  Taking the inverse of the bottom right element of $\boldsymbol{M_{net}}$, we get the transmission $T_{net}$ through the net optical device:
